---
title: "All Calls Exploration of Called Numbers"
subtitle: "Stat 390 Project with Legal Aid Chicago"
author: "Maia Sporn"
format: html
---

```{r}
#| code-fold: true
#| warning: false
# load packages and data
library(tidyverse)
all_calls <- read_csv("all_calls.csv") # this is a cleaned csv
```

Getting some basic summary stats and exploring the issue:

- 344,464 total calls, of these 123,716 have multiple called numbers (36%)

```{r}
#| results: hide
# understand the issue - how many correlation IDs have multiple called numbers
all_calls %>% distinct(correlation_id) %>% nrow() # 344464 total calls

df <- all_calls %>%
  group_by(correlation_id) %>%
  summarize(n_unique_numbers = n_distinct(called_number)) %>%
  filter(n_unique_numbers > 1)
df %>% nrow() # 123716 calls have multiple called numbers (36% of total calls)
```

Exploring hypothesis #1: Transfers

- 57,435 calls with multiple numbers appear to be transfers
  
    - Call transfer time seems to be the key indicator (transfer related call ID is also present for some calls, but always attached to a non-NA call transfer time if present)
    
- Also, 203 calls with only one number are transfers (though this is only a small percentage, 0.9%)

**It doesn't seem like transfers are the sole explanation behind why some calls have multiple called numbers**

```{r}
#| results: hide
# examine relevant variables for calls with multiple numbers
df_leg_info <- all_calls %>%
  group_by(correlation_id) %>%
  filter(n_distinct(called_number) > 1) %>%
  arrange(correlation_id, start_time) %>%
  select(correlation_id, called_number, call_transfer_time, transfer_related_call_id,
         original_reason, related_reason, redirect_reason, related_call_id)
df_leg_info

# compare with calls with one called number
df_one_number <- all_calls %>%
  group_by(correlation_id) %>%
  filter(n_distinct(called_number) == 1) %>%
  arrange(correlation_id, start_time) %>%
  select(correlation_id, direction, called_number, call_transfer_time, releasing_party, 
         original_reason, redirect_reason, related_reason, related_call_id)
df_one_number

df_one_number %>% filter(!is.na(call_transfer_time))

transfer <- df_leg_info %>% filter(!is.na(transfer_related_call_id)) %>%
  select(transfer_related_call_id, correlation_id, called_number, call_transfer_time, 
         original_reason, related_reason, redirect_reason, related_call_id)

calls_with_transfer_id <- unique(transfer$correlation_id)
# 20,151 calls with multiple numbers are attached to a non-NA transfer ID
# but not all call legs with a call transfer time have a transfer ID

call_transfer_time <- df_leg_info %>% filter(!is.na(call_transfer_time)) %>%
  select(transfer_related_call_id, correlation_id, called_number, call_transfer_time,
         original_reason, related_reason, redirect_reason, related_call_id)

calls_with_call_transfer_time <- unique(call_transfer_time$correlation_id)
# 57,435 calls with multiple numbers are attached to a non-NA call transfer time

calls_transfers <- append(calls_with_transfer_id, calls_with_call_transfer_time)
unique(calls_transfers)
# 57,435 calls have call transfer time and/or transfer ID --> 57,435 calls are transfers
# can't be the sole explanation for the 123,716 calls have multiple called numbers 

# find how many calls with one number are transfers
one_number_transfer <- df_one_number %>% filter(!is.na(call_transfer_time))
one_number_calls_with_transfer_id <- unique(one_number_transfer$correlation_id)
# 203 calls with one number are transfers (0.09%)
```

```{r}
# show example of what a transferred call looks like
call_transfer_time %>%
  filter(correlation_id == "9463a568-c2c6-4c11-a536-c87a0fe7a220") %>%
  select(correlation_id, called_number, transfer_related_call_id, call_transfer_time) %>% knitr::kable()
```

Exploring hypothesis #2: Redirection & call routing

- 112,775 calls with multiple numbers have an original/related/redirected reason (again not all 123,716)

- 1728 out of 220,748 calls with only one number have at least one of these routing reasons

**This again suggest redirection & routing is not the source of the multiple-number issue**

```{r}
#| results: hide
# explore redirection / call routing
calls_with_reason <- df_leg_info %>% 
  filter(!is.na(original_reason) | !is.na(related_reason) | !is.na(redirect_reason))
unique(calls_with_reason$correlation_id)
# 112,775 calls have an original/related/redirect reason
# but again not all 123,716

unique(df_one_number$correlation_id)
# 220,748 calls have only one number
# also 1728 of the 220,748 calls with only one number have at least one of these reasons (0.7%)
one_number_calls_with_reason <- df_one_number %>% filter(!is.na(original_reason) | !is.na(related_reason) | !is.na(redirect_reason))
unique(one_number_calls_with_reason$correlation_id)

```

