---
title: "Presentation 4 EDA Checks Data Integrity"
author: "Maia Sporn"
format: html
---

```{r}
#| code-fold: true
#| warning: false
#| output: false
# load packages and data
library(tidyverse)
all_calls <- read_csv("all_calls.csv") # this is a cleaned csv
car <- read_csv("CAR_combined.csv") %>%
  janitor::clean_names()

# find car data range and use it to filter all calls to match
car$activity_start_timestamp_mdy <- as.POSIXct(car$activity_start_timestamp)
all_calls$start_time_mdy <- as.POSIXct(all_calls$start_time)

car_date_min <- min(car$activity_start_timestamp_mdy, na.rm = TRUE)
car_date_max <- max(car$activity_start_timestamp_mdy, na.rm = TRUE)

cat("car date range:", format(car_date_min, "%m/%d"), "-", format(car_date_max, "%m/%d"), "\n")

all_calls_date_min <- min(all_calls$start_time_mdy, na.rm = TRUE)
all_calls_date_max <- max(all_calls$start_time_mdy, na.rm = TRUE)

cat("all_calls date range:", format(all_calls_date_min, "%m/%d"), "-", format(all_calls_date_max, "%m/%d"), "\n")

all_calls_filtered <- all_calls %>% # use this all calls df going forward
  filter(start_time >= car_date_min & start_time <= car_date_max)
```

## Q1

*Is there an internal call that is coming from outside (PSTN vendor should be CallTower and Direction should be ORIGINATING) of Legal Aid? (This should not exist.)*

```{r}
all_calls_filtered %>% 
  filter(pstn_vendor_name == "CallTower" & direction == "ORIGNATING") %>% 
  nrow()
```

This assumption is correct: there are no internal calls that originate from outside of Legal Aid (there are 0 rows with a CallTower PSTN vendor and originating direction).

## Q2

*Do the inbound calls in All Calls data to the 6 target numbers = all calls in CAR? How often do these 6 target numbers appear in the first leg of a call vs. a later leg?*

```{r}
# how often do the 6 numbers of interest appear in the first leg vs a later leg
target_numbers <- c("13123411070", "13124312299", "13125068646", 
                    "13125068647", "13122296080", "13123478342")

call_legs <- all_calls_filtered %>%
  select(correlation_id, called_number, start_time) %>%
  distinct()

call_legs_flagged <- call_legs %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  mutate(is_first = row_number() == 1) %>%
  ungroup()

summary_legs <- call_legs_flagged %>%
  filter(called_number %in% target_numbers) %>%
  mutate(first_status = ifelse(is_first, "is_first", "not_first")) %>%
  count(called_number, first_status) %>%
  pivot_wider(names_from = first_status, values_from = n, values_fill = 0) %>%
  mutate(unique_legs = is_first + not_first)

summary_legs
```

```{r}
# double check the number of unique call legs to the 6 numbers
test <- all_calls_filtered %>% 
  filter(called_number %in% c(
    "13123411070", "13124312299", "13125068646", "13125068647", "13122296080", "13123478342"
    ))

n_distinct(test$correlation_id, test$called_number, test$start_time) 
  # this number should equal the sum of the unique_legs column in above table
```

```{r}
# create a new df to use for testing at the call leg level
leg_level <- all_calls_filtered %>%
  arrange(correlation_id, start_time) %>%
  distinct(start_time, correlation_id, called_number, .keep_all = TRUE) %>%
  group_by(correlation_id) %>%
  mutate(leg_num = row_number()) %>%
  ungroup()
```

**Case 1**: 1st legs only

    - CallTower (external calls only)
    - TERMINATING (inbound)
    - 1 of the 6 numbers appearing in first row/leg 

```{r}
#| label: case-1
## MAIN NUMBER
main <- all_calls_filtered  %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  mutate(row_num = row_number()) %>%
  filter(
    row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13123411070,
    pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
main_calls <- all_calls_filtered %>%
  filter(correlation_id %in% main)
# length(unique(main_calls$correlation_id))

## FARMWORKER
fw <- all_calls_filtered  %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  mutate(row_num = row_number()) %>%
  filter(
    row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13124312299,
    pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
fw_calls <- all_calls_filtered %>%
  filter(correlation_id %in% fw)
# length(unique(fw_calls$correlation_id))

## TRANSFERS
english <- all_calls_filtered %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  mutate(row_num = row_number()) %>%
  filter(
    row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13125068646,
    pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
english_calls <- all_calls_filtered %>%
  filter(correlation_id %in% english)
# length(unique(english_calls$correlation_id))

spanish <- all_calls_filtered %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  mutate(row_num = row_number()) %>%
  filter(
    row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13125068647,
    pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
spanish_calls <- all_calls_filtered %>%
  filter(correlation_id %in% spanish)
# length(unique(spanish_calls$correlation_id))

## OLD ADAPT
adapt <- all_calls_filtered %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  mutate(row_num = row_number()) %>%
  filter(
    row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13122296080,
    pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
adapt_calls <- all_calls_filtered %>%
  filter(correlation_id %in% adapt)
# length(unique(adapt_calls$correlation_id))

## OLD BENEFITS
benefits <- all_calls_filtered %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  mutate(row_num = row_number()) %>%
  filter(
    row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13123478342,
    pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
benefits_calls <- all_calls_filtered %>%
  filter(correlation_id %in% benefits)
# length(unique(benefits_calls$correlation_id))

##

length(unique(main_calls$correlation_id)) + 
  length(unique(fw_calls$correlation_id)) +
  length(unique(english_calls$correlation_id)) +
  length(unique(spanish_calls$correlation_id)) +
  length(unique(adapt_calls$correlation_id)) +
  length(unique(benefits_calls$correlation_id))
```

**Case 2**: 1st legs only except transfers

    - For main number, farmworker, old ADAPT, and old benefits:
    
        - CallTower (external calls only)
        
        - TERMINATING (inbound)
        
        - 1 of the 4 numbers appearing in first row/leg
        
    - For transfers to the English and Spanish main menus:
    
        - PSTN vendor name = NA
        
        - TERMINATING (inbound)
        
        - 1 of the 2 numbers appearing in NOT the first row/leg
        
```{r}
#| label: case-2
## MAIN NUMBER
main <- all_calls_filtered  %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  mutate(row_num = row_number()) %>%
  filter(
    row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13123411070,
    pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
main_calls <- all_calls_filtered %>%
  filter(correlation_id %in% main)
# length(unique(main_calls$correlation_id))

## FARMWORKER
fw <- all_calls_filtered  %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  mutate(row_num = row_number()) %>%
  filter(
    row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13124312299,
    pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
fw_calls <- all_calls_filtered %>%
  filter(correlation_id %in% fw)
# length(unique(fw_calls$correlation_id))

## TRANSFERS
english <- leg_level %>%
  filter(
    leg_num > 1,
    direction == "TERMINATING",
    called_number == 13125068646,
    is.na(pstn_vendor_name)
  ) %>%
  pull(correlation_id)
english_calls <- all_calls_filtered %>%
  filter(correlation_id %in% english)
# length(unique(english_calls$correlation_id))

spanish <- leg_level %>%
  filter(
    leg_num > 1,
    direction == "TERMINATING",
    called_number == 13125068647,
    is.na(pstn_vendor_name)
  ) %>%
  pull(correlation_id)
spanish_calls <- all_calls_filtered %>%
  filter(correlation_id %in% spanish)
# length(unique(spanish_calls$correlation_id))

## OLD ADAPT
adapt <- all_calls_filtered %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  mutate(row_num = row_number()) %>%
  filter(
    row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13122296080,
    pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
adapt_calls <- all_calls_filtered %>%
  filter(correlation_id %in% adapt)
# length(unique(adapt_calls$correlation_id))

## OLD BENEFITS
benefits <- all_calls_filtered %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  mutate(row_num = row_number()) %>%
  filter(
    row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13123478342,
    pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
benefits_calls <- all_calls_filtered %>%
  filter(correlation_id %in% benefits)
# length(unique(benefits_calls$correlation_id))

##

length(unique(main_calls$correlation_id)) + 
  length(unique(fw_calls$correlation_id)) +
  length(unique(english_calls$correlation_id)) +
  length(unique(spanish_calls$correlation_id)) +
  length(unique(adapt_calls$correlation_id)) +
  length(unique(benefits_calls$correlation_id))
```

**Case 3**: Any leg, any source

    - No filter on PSTN vendor name
    
    - TERMINATING (inbound)
        
    - 1 of the 6 numbers appearing in any row/leg
    
```{r}
#| label: case-3
## MAIN NUMBER
main <- all_calls_filtered  %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  # mutate(row_num = row_number()) %>%
  filter(
    # row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13123411070,
    # pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
main_calls <- all_calls_filtered %>%
  filter(correlation_id %in% main)
# length(unique(main_calls$correlation_id))

## FARMWORKER
fw <- all_calls_filtered  %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  # mutate(row_num = row_number()) %>%
  filter(
    # row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13124312299,
    # pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
fw_calls <- all_calls_filtered %>%
  filter(correlation_id %in% fw)
# length(unique(fw_calls$correlation_id))

## TRANSFERS
english <- leg_level %>%
  filter(
    # leg_num > 1,
    direction == "TERMINATING",
    called_number == 13125068646,
    # is.na(pstn_vendor_name)
  ) %>%
  pull(correlation_id)
english_calls <- all_calls_filtered %>%
  filter(correlation_id %in% english)
# length(unique(english_calls$correlation_id))

spanish <- leg_level %>%
  filter(
    # leg_num > 1,
    direction == "TERMINATING",
    called_number == 13125068647,
    # is.na(pstn_vendor_name)
  ) %>%
  pull(correlation_id)
spanish_calls <- all_calls_filtered %>%
  filter(correlation_id %in% spanish)
# length(unique(spanish_calls$correlation_id))

## OLD ADAPT
adapt <- all_calls_filtered %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  # mutate(row_num = row_number()) %>%
  filter(
    # row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13122296080,
    # pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
adapt_calls <- all_calls_filtered %>%
  filter(correlation_id %in% adapt)
# length(unique(adapt_calls$correlation_id))

## OLD BENEFITS
benefits <- all_calls_filtered %>%
  arrange(correlation_id, start_time) %>%
  group_by(correlation_id) %>%
  # mutate(row_num = row_number()) %>%
  filter(
    # row_num == 1, 
    direction == "TERMINATING", 
    called_number == 13123478342,
    # pstn_vendor_name == "CallTower"
  ) %>%
  pull(correlation_id)
benefits_calls <- all_calls_filtered %>%
  filter(correlation_id %in% benefits)
# length(unique(benefits_calls$correlation_id))

##

length(unique(main_calls$correlation_id)) + 
  length(unique(fw_calls$correlation_id)) +
  length(unique(english_calls$correlation_id)) +
  length(unique(spanish_calls$correlation_id)) +
  length(unique(adapt_calls$correlation_id)) +
  length(unique(benefits_calls$correlation_id))
```

None of these three cases seem to quite add up to the number of distinct contact session IDs in CAR. Different filters to test? 
        
## Q3

*Internal transfers to individual categories in All Calls data = Calls to those categories in CAR?*

Checking to see if farmworker calls overlap across the two datasets:

```{r}
# farmworker in car
fw_all_calls <- all_calls_filtered %>% 
  filter(called_number == "13124312299" & 
           direction=="TERMINATING" & 
           is.na(pstn_vendor_name))
length(unique(fw_all_calls$correlation_id))

# first test EP name
fw_ep <- car %>% 
  filter(ep_name == "Farmworker Main Number Telephony EP")
length(unique(fw_ep$contact_session_id)) # more than what appears in all calls

# try adding filters to reduce the number counted in car
fw_ep_activity <- car %>% 
  filter(
    ep_name == "Farmworker Main Number Telephony EP" &
      activity_name == "FarmworkerMainMenu"
  )
length(unique(fw_ep_activity$contact_session_id)) # still greater than 6025

# add another filter
fw_ep_activity_flow <- car %>% 
  filter(
    ep_name == "Farmworker Main Number Telephony EP" &
      activity_name == "FarmworkerMainMenu" &
      flow_name == "FarmworkerMain"
  )
length(unique(fw_ep_activity_flow$contact_session_id))

# remove a filter
fw_ep_flow <- car %>% 
  filter(
    ep_name == "Farmworker Main Number Telephony EP" &
      flow_name == "FarmworkerMain"
  )
length(unique(fw_ep_flow$contact_session_id))

# look at internal transfers only for all calls
fw_all_calls <- leg_level %>% 
  filter(leg_num > 1, called_number == "13124312299", 
         direction=="TERMINATING", 
           is.na(pstn_vendor_name))
length(unique(fw_all_calls$correlation_id)) # none
```

Other categories to look at?

## Q4
*How many internal transfers vs direct calls?*

```{r}
call_legs <- all_calls_filtered %>%
  filter(direction == "TERMINATING") %>%
  arrange(correlation_id, start_time) %>%
  distinct(correlation_id, called_number, start_time, .keep_all = TRUE)

nrow_terminating <- nrow(call_legs)
nrow_terminating

nrow_direct <- call_legs %>%
  filter(pstn_vendor_name == "CallTower") %>%
  nrow()
nrow_direct

nrow_transfers <- call_legs %>%
  filter(is.na(pstn_vendor_name)) %>%
  nrow()
nrow_transfers

# nrow_transfers + nrow_direct should equal nrow_terminating

nrow_direct + nrow_transfers
```

## Q5

*Are there calls of zero duration in CAR?*

```{r}
#| warning: false
library(lubridate)

# at the contact session ID level
car_adj <- car %>%
  mutate(
    activity_start_timestamp_clean = parse_date_time(
      activity_start_timestamp, 
      orders = c("ymd HMS p", "ymd HM p", "ymd H p"))
  )

call_durations <- car_adj %>%
  group_by(contact_session_id) %>%
  summarize(
    num_rows = n(),
    start_time = min(activity_start_timestamp_clean),
    end_time = max(activity_start_timestamp_clean),
    duration_secs = as.numeric(difftime(end_time, start_time, units = "secs"))
  ) %>%
  ungroup()

zero_duration_calls <- call_durations %>%
  filter(duration_secs == 0)

single_row_calls <- call_durations %>%
  filter(num_rows == 1)

nrow(zero_duration_calls)
nrow(single_row_calls)
length(unique(zero_duration_calls$contact_session_id)) +
  length(unique(single_row_calls$contact_session_id))

zero_all_calls <- all_calls_filtered %>% filter(duration==0)
length(unique(zero_all_calls$correlation_id))

```

In All Calls, 29,836 correlation IDs contain a leg with 0 duration. However, in CAR, 1,228 contact session IDs have 0 seconds between activities and 1 contact session ID only has a single row: 1,229 contact session IDs with 0 duration.


## Q6

*What other popular numbers are there, besides special intake lines, and main number, and farmworker number?*

```{r}
inbound_legs <- all_calls_filtered %>%
  filter(direction == "TERMINATING") %>%
  select(correlation_id, called_number, start_time) %>%
  distinct()

inbound_counts <- inbound_legs %>%
  group_by(called_number) %>%
  tally(name = "num_legs") %>%
  arrange(desc(num_legs))

inbound_counts <- inbound_counts %>%
  mutate(
    description = case_when(
      called_number == "13123411070" ~ "Main Number",
      called_number == "13124312299" ~ "Farmworker",
      called_number == "13125068646" ~ "English Main Menu",
      called_number == "13125068647" ~ "Spanish Main Menu",
      called_number == "13122296080" ~ "Old ADAPT",
      called_number == "13123478342" ~ "Old Benefits",
      called_number == "13123478347" ~ "VM for A2J immigration",
      called_number == "18882652188" ~ "A2J immigration toll free",
      called_number == "13124235904" ~ "Austin Intake VM",
      called_number == "13122296344" ~ "Bankruptcy Helpdesk VM",
      called_number == "13124235900" ~ "CLASP VM",
      called_number == "13122296071" ~ "Criminal Records",
      called_number == "13123478392" ~ "Education Law Referrals VM",
      called_number == "13124235909" ~ "Fair Housing Intake VM",
      called_number == "13123478309" ~ "HIV Intake VM",
      called_number == "13122296072" ~ "JEHD (Juvenile Expungement Help Desk)",
      called_number == "13124235938" ~ "legalclinics (Community Legal Clinics)",
      called_number == "13124312101" ~ "OP Appeals Project",
      called_number == "13123478340" ~ "Veterans Rights Project VM",
      called_number == "13122296073" ~ "Trafficking Survivors Assistance Project",
      called_number == "18004459025" ~ "Migrant Legal Assistance Program",
      called_number == "18884018200" ~ "Nursing Home Ombudsman",
      called_number == "13122296014" ~ "Markham Eviction Help Desk",
      TRUE ~ NA
    )
  )
inbound_counts
```

## Q7

*How many calls have 1 leg?*

```{r}
# define a unique call leg using distinct (start_time, correlation_id, called_number)
call_legs_new <- all_calls_filtered %>%
  select(start_time, correlation_id, called_number) %>%
  distinct()

leg_counts <- call_legs_new %>%
  group_by(correlation_id) %>%
  tally(name = "num_legs")

num_single_leg_calls <- leg_counts %>%
  filter(num_legs == 1) %>%
  summarise(count = n()) %>%
  pull(count)

num_single_leg_calls
```

209,907 calls have one leg.


